On The Difficulty of Computing Non-Commutative Signature Gröbner Bases

### 2022-07-15 article_issac.tex
section 7

Let $(f_1,f_2)=(x-y,y^2) \in \mathbb{Q} \left\langle x,y \right\rangle^2,$
equipped with the \textit{deglex} monomial ordering $\leq.$
We prove that $Syz(f_1,f_2)$ admits a finite GB 
for the following module ordering: for $a_1 \varepsilon_{i_1} b_1, a_2 \varepsilon_{i_2} b_2 \in \MM \left( \mathscr{F}_2 \right)$, we say that  $a_1 \varepsilon_{i_1} b_1 \leq a_2 \varepsilon_{i_2} b_2$ if:
\begin{enumerate}
\item $a_1<a_2$.
\item $a_1=a_2$ and $b_1 < b_2,$
\item $a_1=a_2,$ $b_1 = b_2,$ and $i_1 \leq i_2.$
\end{enumerate}

We then have the following fact:
\begin{itemize}
\item $s_1=\underline{xy \varepsilon_2}-\varepsilon_2 y^2+ x \varepsilon_1 y^2-y \varepsilon_1 y^2- \varepsilon_1 x y^2 \in Syz(f_1,f_2)$,
\item $s_2=\underline{xy \varepsilon_1}-yx\varepsilon_1 -  \varepsilon_1 x^2+y \varepsilon_1 xy \in Syz(f_1,f_2)$.
\end{itemize}
In addition, the following are trivial syzygys:

\begin{itemize}
\item $s_3=\underline{x \varepsilon_1}-y\varepsilon_1- \varepsilon_1 x + \varepsilon_1 y \in Syz(f_1,f_2)$,
\item $s_4=\underline{y^2 \varepsilon_1}-\varepsilon_2 x +  \varepsilon_2 y \in Syz(f_1,f_2)$.
\item $s_5=\underline{x \varepsilon_2}-y\varepsilon_2- y^2 \varepsilon_2  \in Syz(f_1,f_2)$,
\item $s_6=\underline{y^2 \varepsilon_2}-\varepsilon_2 y^2 \in Syz(f_1,f_2)$.
\end{itemize}

We then state that:
\begin{lem}
The module of syzygys $Syz(f_1,f_2)$ admits $(s_1,\dots,s_6)$
as a Gröbner basis.
\end{lem}
\begin{proof}
Let $m=a \varepsilon_1 b \in \MM \left( \mathscr{F}_2 \right)$
be a leading monomial of a syzygy $s$.
Then we distinguish the following cases.
\begin{enumerate}
\item If $a = \dots x$ then $m$ is divisible by $\LM(s_3).$
\item If $a = \dots xy$ then $m$ is divisible by $\LM(s_2).$
\item If $a = \dots y^2$ then $m$ is divisible by $\LM(s_4).$
\item If $a = y,$ then $s = y \varepsilon_1 u+  \varepsilon_1 v + \varepsilon_2 w$ for some polynomials $u,v,w.$
Applying $\overline{\cdot}$ (mapping $\varepsilon_i$ to $f_i$), we 
obtain that $y (x-y) u+  (x-y) v + y^2 w=0.$
On the left hand-side, the polynomial $xv$ contains all the monomials
with prefix $x$, which implies 
that no cancellation with other terms can occur,
and then $v=0.$
Similarly, $yxu$ contains all the monomials
with prefix $yx$, which then implies that
$u=0$ and then $y^2 w=0$ and $w=0$, and we get
a contradiction.
\item If $a = 1,$ then $s = \varepsilon_1 u+  \varepsilon_2 v $ for some polynomials $u,v.$
Again, applying $\overline{\cdot}$, we 
obtain that $u=v=0$ which is a contradiction.
\end{enumerate}
Similarly, let $m=a \varepsilon_2 b \in \MM \left( \mathscr{F}_2 \right)$
be a leading monomial of a syzygy $s$. Then:
\begin{enumerate}
\item If $a = \dots x$ then $m$ is divisible by $\LM(s_5).$
\item If $a = \dots xy$ then $m$ is divisible by $\LM(s_1).$
\item If $a = \dots y^2$ then $m$ is divisible by $\LM(s_6).$
\item If $a = y,$ we obtain a contradiction.
\item If $a = 1,$ we also obtain a contradiction.
\end{enumerate}

>>>>

Let $(f_1,f_2)=(x-y,y^2) \in \mathbb{Q} \left\langle x,y \right\rangle^2,$
equipped with the \textit{deglex} monomial ordering $\leq.$
We prove that $Syz(f_1,f_2)$ admits a finite GB 
for the following module ordering: for $a_1 \varepsilon_{i_1} b_1, a_2
\varepsilon_{i_2} b_2 \in \MM \left( \mathscr{F}_2 \right)$, we say that  $a_1
\varepsilon_{i_1} b_1 < a_2 \varepsilon_{i_2} b_2$ if one of the following
conditions holds:
\begin{enumerate}
\item $a_1<a_2$.
\item $a_1=a_2$ and $b_1 < b_2,$
\item $a_1=a_2,$ $b_1 = b_2,$ and $i_1 < i_2.$
\end{enumerate}

We then have the following fact:
\begin{itemize}
\item $s_1=\underline{xy \varepsilon_2}-\varepsilon_2 y^2+ x \varepsilon_1 y^2-y \varepsilon_1 y^2- \varepsilon_1 x y^2 \in Syz(f_1,f_2)$,
\item $s_2=\underline{xy \varepsilon_1}-yx\varepsilon_1 -  \varepsilon_1 x^2+y
  \varepsilon_1 xy \in Syz(f_1,f_2)$.
\fixmecc{j'ai l'impression que ce n'est pas une syzygie. On peut prendre à la
  place $s_2=\underline{xy\varepsilon_1}-\varepsilon_1yx+x\varepsilon_2-\varepsilon_2x$}  
\end{itemize}
In addition, the following are trivial syzygys:

\begin{itemize}
\item $s_3=\underline{x \varepsilon_1}-y\varepsilon_1- \varepsilon_1 x + \varepsilon_1 y \in Syz(f_1,f_2)$,
\item $s_4=\underline{y^2 \varepsilon_1}-\varepsilon_2 x +  \varepsilon_2 y \in Syz(f_1,f_2)$.
\item $s_5=\underline{x \varepsilon_2}-y\varepsilon_2- y^2 \varepsilon_2  \in
  Syz(f_1,f_2)$,
  \fixmecc{ce ne serait pas plutôt
    $\underline{x\varepsilon_2}-y\varepsilon_2-\varepsilon_1y^2$?} 
\item $s_6=\underline{y^2 \varepsilon_2}-\varepsilon_2 y^2 \in Syz(f_1,f_2)$.
\end{itemize}

We then state that:
\begin{lem}
The module of syzygys $Syz(f_1,f_2)$ admits $(s_1,\dots,s_6)$
as a Gröbner basis.
\end{lem}
\begin{proof}
Let $m=a \varepsilon_1 b \in \MM \left( \mathscr{F}_2 \right)$
be a leading monomial of a syzygy $s$.
Then we distinguish the following cases.
\begin{enumerate}
\item If $a = \dots x$ then $m$ is divisible by $\LM(s_3).$
\item If $a = \dots xy$ then $m$ is divisible by $\LM(s_2).$
\item If $a = \dots y^2$ then $m$ is divisible by $\LM(s_4).$
\item If $a = y,$ then $s = y \varepsilon_1 u+  \varepsilon_1 v + \varepsilon_2 w$ for some polynomials $u,v,w.$
Applying $\overline{\cdot}$ (mapping $\varepsilon_i$ to $f_i$), we 
obtain that $y (x-y) u+  (x-y) v + y^2 w=0.$
On the left hand-side, the polynomial $xv$ contains all the monomials
with prefix $x$, which implies 
that no cancellation with other terms can occur,
and then $v=0.$
Similarly, $yxu$ contains all the monomials
with prefix $yx$, which then implies that
$u=0$ and then $y^2 w=0$ and $w=0$, and we get
a contradiction.
\item If $a = 1,$ then $s = \varepsilon_1 u+  \varepsilon_2 v $ for some polynomials $u,v.$
Again, applying $\overline{\cdot}$, we 
obtain that $u=v=0$ which is a contradiction.
\end{enumerate}
Similarly, let $m=a \varepsilon_2 b \in \MM \left( \mathscr{F}_2 \right)$
be a leading monomial of a syzygy $s$. Then:
\begin{enumerate}
\item If $a = \dots x$ then $m$ is divisible by $\LM(s_5).$
\item If $a = \dots xy$ then $m$ is divisible by $\LM(s_1).$
\item If $a = \dots y^2$ then $m$ is divisible by $\LM(s_6).$
\item If $a = y,$ we obtain a contradiction \fixmecc{on peut avoir
  $\underline{y\varepsilon_2}-\varepsilon_2y$, mais c'est une syzygie principale}.
\item If $a = 1,$ we also obtain a contradiction.
\end{enumerate}
